### 2020-09-13 Thesis-Chapter1-Introduction.tex
Additional reading. 'Wicked Crypto' is an excellent source
Address some TODOs, fill in judicial role

Difficult policy matters are \acp{wicked-prob}. Originally coming from the field of design theory, the term
``\ac{wicked-prob}'' uses ``wicked'' not in the moral sense, but in the malignant, vicious, and tricky sense
\cite{rittel_dilemmas_1973}. Unlike their ``tame'' counterparts which science and engineering are well equipped to
answer, \acp{wicked-prob} lack clear formulations, causes, resolutions, or measurements. Each attempted solution has
permanent and often unintended consequences, and likely exists in a pattern of chronic policy failure.

\Acl{EA} is a wicked problem. In encryption policy, \ac{EA} is alternative means of decryption intended for law
enforcement use. Characterized by a dynamic technological environment, disagreement about underlying values, and
resistance to a clear solution, the debate on \ac{EA} refuses to end. This thesis will not attempt to end the debate,
but to inform it with additional structure.

>>>>

Certain policy issues are \acp{wicked-prob}. Originally coming from the field of design theory, the term
``\ac{wicked-prob}'' uses ``wicked'' not in the moral sense, but in the malignant, vicious, and tricky sense
\cite{rittel_dilemmas_1973}. Unlike their ``tame'' counterparts which science and engineering are well equipped to
answer, \acp{wicked-prob} lack clear formulations, causes, resolutions, or measurements. Each attempted solution has
permanent and often unintended consequences, and likely exists in a pattern of chronic policy failure.

\Acl{EA} is a wicked problem \cite{rozenshtein_wicked_2018}. In encryption policy, \ac{EA} is alternative means of
decryption intended for law enforcement use. Characterized by a dynamic technological environment, disagreement about
underlying values, and resistance to a clear solution, the debate on \ac{EA} will not go away. This thesis does not
attempt to end the debate, but to structure and analyze it.

### 2020-09-13 Thesis-Chapter2-Background.tex
Improve argdown map generation build process
Create Latex macro for argument map includes
Address some TODOs, fill in judicial role

Data secrecy is not cryptography's only purpose, however.

Cryptography is the technical foundation for many forms of computer and network security. Security is defined in terms
of several properties: authentication, integrity, non-repudiation, confidentiality, availability, and authorization
\cite{shostack_threat_2014}. These system-level properties emerge from both the composition of the system's components
(architecture) and the security of the components themselves; likewise, the security of each component emerges from the
architecture of its subcomponents and the security of the subcompenents themselves. At the bottom of this analysis, the
security of primitive components commonly relies on cryptography.

@@

% TODO: clarify relationship between encryption and cryptography. Use appropriately throughout.

@@

\subsubsection{Enacted U.S. Regulations}
% FIXME: distinguish statutes vs. regulations; rename "failed" section

@@

Amended the Wiretap Act, Stored Communications Act, and Pen Register Act. Provided explicit protections for private
% FIXME: Briefly introduce these. Even just "Amended the switch-board-era"...
electronic communications data, and updated definitions to contemporary technology. It granted law enforcement explicit
access to certain data as well; some terminology has aged poorly with technological evolution, making the access
increasingly permissive \cite{shamsi_2011}.

@@

The above history and section on U.S. law encompasses the executive and legislative branches' influence on surveillance
and cryptography. A complete picture requires understanding the judiciary branch's influence via court rulings. The
judiciary branch exercises power through a framework based on AAAA and BBBB. In this section we analyze several
important cases and developments in the last few years.

% TODO: Framework for understanding judicial influence. See bergerson_judicial_1998 and gizzi_fourth_2016. Update the
%   "AAAA and BBBB" in the intro.

% TODO: Review historically important cases on the 1st, 4th, and 5th amendments, e.g. Katz v. United States, 1969.

% TODO: Review recent cases. See orenstein_2016, bittenbender_2019, vaas_2019.
% NOTE: _Carpenter_ was a recent important Supreme Court case limiting the scope of the third-party doctrine!
% See franklin_2018 and rozenshtein_2018

\subsubsection{Failed U.S. Regulation Attempts}
\label{sec-us-reg-fails}

@@

@@

This thesis is U.S.--focused, but \ac{EA} technologies and policies have international impacts. The following is a very
brief account of the regulatory environment in various jurisdictions around the world based primarily on a 2016 analysis
by the Law Library of Congress \cite{acosta_government_2016}.

@@

France, Belgium, Germany, and Sweden require cooperation between telecommunications providers and law enforcement; for
France, Belgium, and Germany, that includes decrypting network traffic when possible. France, Germany, and Sweden each
have some level of \ac{lawful-hacking} capability. There is no E.U. legislation that requires key disclosure or
decryption of network traffic.

@@

Japanese authorities require cooperation and can request access to decrypted data, but subjects are not punished for
declining such requests. Taiwanese regulation does not specifically address encryption, but does mandate that
telecommunications providers provide interfaces ``with functions that can cooperate with interception.'' Chinese
authorities have great access to plaintext data by virtue of the political structure of the state, but its 2015
counter-terrorism law explicitly compels access to plaintext data upon request \cite{donahue_comparative_2018}.

@@

% TODO: add wright_crypto_2018 (I _think_ it goes here)

\propsend

@@

    \prop{Forced password disclosure}{

In some cases, a suspect in custody may have the ability to access the desired encrypted information via a key,
\ac{PIN}, or password. In these cases, authorities sometimes compel the suspect to disclose the password. In the U.S.,
this approach draws debates about the fifth amendment right against self-incrimination \cite{bittenbender_2019}. Of
course, this approach cannot address \ac{DIM} or situations in which the knowledgeable party is not in custody.

@@

\begin{figure}[ht]
    \centering\CaptionFontSize
    \includegraphics[width=\linewidth]{arguments/build/demo.pdf}
    \caption[Argdown Demo]{A demonstrative example of argument maps with Argdown.}
    \label{fig-args-demo}
\end{figure}

>>>>

Data secrecy is not cryptography's only use, however. Precisely speaking, data secrecy is the purpose of encryption.
Encryption is cryptography's ``killer app,'' but it is only a part of cryptography's usefulness to security.

Cryptography is the technical foundation for many forms of computer and network security. Security is defined in terms
of several properties: authentication, integrity, non-repudiation, confidentiality, availability, and authorization
\cite{shostack_threat_2014}. These system-level properties emerge from both the composition of the system's components
(architecture) and the security of the components themselves; likewise, the security of each component emerges from the
architecture of its subcomponents and the security of the subcompenents themselves. At the end of this chain of
analysis, the security of primitive components commonly relies on cryptography.

@@

@@

\subsubsection{Enacted U.S. Regulations}

@@

Amended the Wiretap Act, Stored Communications Act, and Pen Register Act. Provided explicit protections for private
% TODO: Briefly introduce these. Even just "Amended the switch-board-era"...
electronic communications data, and updated definitions to contemporary technology. It granted law enforcement explicit
access to certain data as well; some terminology has aged poorly with technological evolution, making the access
increasingly permissive \cite{shamsi_2011}.

@@

\subsubsection{Unsuccessful U.S. Regulation Attempts}
\label{sec-us-reg-fails}

@@

The above history and section on U.S. law encompasses the executive and legislative branches' influence on surveillance
and cryptography. A complete picture requires understanding the judiciary branch's influence via court rulings. The
judiciary branch exercises power through a framework based on AAAA and BBBB. In this section I describe this framework,
review how it has influenced cryptography policy, and analyze where cryptographic jurisprudence may go from here.

% TODO: Framework for understanding judicial influence. See bergerson_judicial_1998. Update the "AAAA and BBBB" in the
% intro and the todo in the conclusion.

Judicial influence on encryption policy is primarily mediated through Fourth Amendment jurisprudence, particularly
regarding the interpretation of a person's ``effects,'' what is ``reasonable,'' and what is a ``search.'' The Supreme
Court's interpretation and implementation of the Amendment's crime control-- and civil liberty--balancing principles has
changed over time. An in-depth analysis of the Court's jurisprudence from the 1960's Warren Court, through the
influential 1980s Rehnquist Court, to the mid-2010s Roberts Court found a deep proclivity towards crime control that is
slowly thawing \cite{gizzi_fourth_2016}. This thawing, caused by evolution of technology, of the court, and of the
justices themselves, has been favorable for data privacy. Recent cases have increasingly sided with defendants, have
caused justices to strain to show that they understand technology, and have increased the categories of protected
digital data \cite{gizzi_fourth_2016}. 2018's potentially landmark \ii{Carpenter v. United States} may be the most
impactful yet; though it only resulted in a narrow decision against unwarranted cell phone location data tracking, it
does undermine the pre-digital third-party doctrine, which holds that Fourth Amendment protection is forfeit when the
data is willingly given to a third-party \cite{franklin_2018}.

As this relates to \ac{EA}, Fourth Amendment interpretation practically determines the threshold at which authorities
require a warrant, not what demands they can or cannot make once they have one. The second most salient legal issue is
the Fifth Amendment right against self-incrimination, which comes up in cases of compelled password disclosure. This
question has not yet reached the Supreme Court, and state courts have ruled compelled disclosure both constitutional
\cite{sobel_2019} \cite{lee_nj_2020} and unconstitutional \cite{lee_its_2020} \cite{vaas_2019}.

% Paragraph interpreting this change as principles vs rules in a wicked problem

That judicial influence on privacy and technology is changing is to be expected. To explain the fluctuating state of
legal interpretation as it relates to \ac{EA}, Matyas et al. make the helpful distinction between legal rules and
principles \cite{matyas_incommensurability_2018}:

\begin{displayquote}
The tension between technical and legal views of sensitive issues such as encryption and surveillance is illuminated by
applying the jurisprudential lens. \dots~Liberal legal systems, manifesting what is generally understood as ``the rule
of law,'' are actually composed of both rules and principles. Legal ``rules'' can be understood as logical propositions
that are expected to yield answers about what is and is not permitted using formal reasoning capabilities. By contrast,
legal ``principles'' articulate values and policies that must be reflected in a legal system but do not necessarily
dictate an unambiguous outcome in any given case.
\end{displayquote}

Rules are the low-level implementation of principles. When the underlying facts of the situation change, as they do by
definition for wicked problems, the rules become out of sync with the principles that bade them. Recent flux in judicial
rulings surrounding surveillance and technology are evidence that the courts are in the process of updating
interpretations of rules in light of the new facts; which principles they ultimately choose to favor remains to be seen,
but rulings like \ii{Carpenter} show a willingness to protect privacy.
% TODO: Connect back to system of judicial influence--this willingness uses that framework to support encryption.

@@

This thesis is U.S.--focused, but \ac{EA} technologies and policies have international impacts. The following is a very
brief account of the regulatory environment in various geographical jurisdictions around the world based primarily on a
2016 analysis by the Law Library of Congress \cite{acosta_government_2016}. Globally, a 2017 study found that more than
half the world's internet users lived in a country where some form of plaintext recovery is mandated \cite{lewis_2017}.

@@

Among European countriues, France and Russia join the U.K. in having access mandates \cite{lewis_2017}. France, Belgium,
Germany, and Sweden require cooperation between telecommunications providers and law enforcement; for France, Belgium,
and Germany, that includes decrypting network traffic when possible. France, Germany, and Sweden each have some level of
\ac{lawful-hacking} capability. There is no E.U. legislation that requires key disclosure or decryption of network
traffic.

@@

Japanese authorities require cooperation and can request access to decrypted data, but subjects are not punished for
declining such requests. Taiwanese regulation does not specifically address encryption, but does mandate that
telecommunications providers provide interfaces ``with functions that can cooperate with interception.'' Chinese
authorities have great access to plaintext data by virtue of the political structure of the state, but also through
explicit requirements in recent anti-terrorism and cybersecurity laws \cite{lewis_2017}.

@@

    \prop{Cryptographic ``crumple zones''}{

A recent proposal adopts the idea of crumple zones from automotive engineering---``in an emergency situation the
construction should break a little bit in order to protect the integrity of the system as a whole and the safety of its
human users'' \cite{wright_crypto_2018}. Similar to the translucent approach, no keys are escrowed and only passive
surveillance is possible. Strong cryptography is a precondition for this scheme, in which keys are generated in a manner
that makes them inherently recoverable only after a large up front and marginal cost.

@@

    \prop{Compelled password disclosure}{

In some cases, a suspect in custody may have the ability to access the desired encrypted information via a key,
\ac{PIN}, or password. In these cases, authorities sometimes compel the suspect to disclose the password. In the U.S.,
this approach draws debates about the fifth amendment right against self-incrimination \cite{bittenbender_2019}
\cite{sobel_2019}. Of course, this approach cannot address \ac{DIM} or situations in which the knowledgeable party is
not in custody.

@@

\begin{figure}[ht]
    \centering\CaptionFontSize
    \myincludeargument{demo}
    \caption[Argdown Demo]{A demonstrative example of argument maps with Argdown.}
    \label{fig-args-demo}
\end{figure}
